---
title: "LAEI River data"
author: "James Smith"
output: html_document
date: "`r format(Sys.time(), '%Y-%m%-%d %H%:%M')`"
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "docs") }) 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(sf)
library(mapview)
library(leaflet)
library(tidyverse)
library(BAMMtools)
library(raster)
library(sp)
```

```{r, echo=F}
latlong = "+init=epsg:4326"
ukgrid = "+init=epsg:27700"
google = "+init=epsg:3857"
```

```{r import data, include=FALSE}

#berths <- st_read('X:/Projects/PLA/GIS/Shapefiles/Berths.shp') %>% st_set_crs(27700) %>% st_transform(4326)
#berths <- berths[-179,]
#river  <- st_read('X:/Projects/PLA/GIS/Shapefiles/River Thames Outline.shp') %>% st_set_crs(27700) %>% st_transform(4326)
```

## Map of berths and river shapefile

```{r plot, echo=F}
#mapview(berths, map.types = c("OpenStreetMap.BlackAndWhite")) %>% addFeatures(river)
```

## 100m grid of Thames emissions

```{r grid_emissions, echo=F}

emissions                   <- read_csv('emissions/inventory_export_2016.csv', col_types = cols())
emissions                   <- aggregate(data=emissions, cbind(Sailing_kg, AtBerth_kg) ~ Substance + CellID, FUN=sum)
emissions$CellID            <- as.numeric(emissions$CellID)

grid                        <- st_read('grids/LAEIGridExtensionV2.gpkg', quiet = T)

grid_emissions              <- merge(emissions, grid, by.x = 'CellID', by.y = 'GRID_ID') 
st_geometry(grid_emissions) <- grid_emissions$geom
```

Example of NOX while sailing

```{r sailing_nox, echo=F}
NOX <- grid_emissions[grid_emissions$Substance == 'NOx',]
```

```{r, echo=F}
mapview(NOX, map.types = c("OpenStreetMap.BlackAndWhite"), zcol = 'Sailing_kg', at = round(getJenksBreaks(NOX$Sailing_kg, 10),0))

rm(NOX)
```

Example of NOX while at berth

```{r berth_nox, echo=F}
NOX <- grid_emissions[grid_emissions$Substance == 'NOx',]
```

```{r, echo=F}
mapview(NOX, map.types = c("OpenStreetMap.BlackAndWhite"), zcol = 'AtBerth_kg', at = round(getJenksBreaks(NOX$AtBerth_kg, 10),0))

rm(NOX)
```

Now make a 10m x 10m grid covering the area of the river emissions. Filled with dummy data for now to visualise.

```{r make_10m_grid, echo=FALSE, warning=FALSE}

new_grid                  <- raster(extent(grid_emissions), res=10, crs = ukgrid)
new_grid$sailing_points   <- NA
```

```{r}
mapview(new_grid$sailing_points, map.types = c("OpenStreetMap.BlackAndWhite"))
```

Now import one day of points from the processed AIS data (via Andrew), and count the number of points within each 10m cell

```{r import_ais_data}

load('gps/apr1_AIS_decoded.Rdata')
gps_data <- data
rm(data)
coordinates(gps_data) <- ~lon+lat
proj4string(gps_data)  <- CRS(latlong)
gps_data <- spTransform(gps_data, CRS(ukgrid))

gps_data <- crop(gps_data, new_grid)

gps_data$time <- 1

gps_grid <- rasterize(gps_data, new_grid, fun=sum)

names(gps_grid) <- c('id', 'count')

```

```{r}
mapview(new_grid$count, map.types = c("OpenStreetMap.BlackAndWhite"))

```

Aggregate the 10m count up to 100m so that know the number of points in each 100m cell

```{r aggregate_up_to_100m}

hundred_grid <- aggregate(gps_grid, fact = 10, fun=sum)

```